Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2020-10-18 GWLDCI.tex
addressed reference comments

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$ we can consider the restriction $C(W)|_V$ of $C(W)$, which consists of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, if $C(W)|_V$ has fewer than $|V|$ rows of non-zero entries then it does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, then $V$ is a dependent set. Any set of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$. As mentioned above, Theorem [3.38] of \cite{wilsonloops} shows that if $W$ is weakly admissible, the matroid $M(W)$ is a positroid. Furthermore, Theorem 5.1 \cite{basisshapeloci} shows that the space parametrized by the matrices $C(W)$ matches the positroid cells defined by $M(W)$, up to a set of measure $0$. Note, this does not imply that every possible specializaton of $C(W)$ to a real valued matrix leads to a postroid, only that the set of values that do define a positroid is volume filling in $\R^{4k}$. \todo{What is [6] referred to by reviewer M?}

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,[n])$ be a weakly admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \ \textrm{ such that } \ \sqcup P_i = \cP; \   \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item The set $F(P)$ from Definition~\ref{VPropdfn} is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\item If $F$ is a cyclic flat of $M(W)$, then it is a propagator flat.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of item 1 is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. Item 2 is \cite[Lemma 3.27]{wilsonloop}. Item 3 is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$, and item 5 is given in \cite[Lemma 3.28]{wilsonloop}.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat show that certain subdiagrams of $W$ can be realized as contractions of $M(W)$:

\begin{lem} \label{contractsubdiaglem} \cite[Corollary 3.33]{wilsonloop} 
Let $W = (\cP, [n])$ be a Wilson loop diagram and $P \subseteq \cP$. If the set $V(P)^c$ has rank $|P^c|$, then the matroid defined by the subdiagram $(P, V(P))$ is equal to the contraction $M(W)/V(P)^c$.
\end{lem}

>>>>

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$ we can consider the restriction $C(W)|_V$ of $C(W)$, which consists of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, if $C(W)|_V$ has fewer than $|V|$ rows of non-zero entries then it does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, then $V$ is a dependent set. Any set of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$. As mentioned above, Theorem [3.38] of \cite{wilsonloop} shows that if $W$ is weakly admissible, the matroid $M(W)$ is a positroid. Furthermore, Theorem 5.1 \cite{WLDdim} shows that the space parametrized by the matrices $C(W)$ matches the positroid cells defined by $M(W)$, up to a set of measure $0$. Note, this does not imply that every possible specializaton of $C(W)$ to a real valued matrix leads to a postroid, only that the set of values that do define a positroid is volume filling in $\R^{4k}$.

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,[n])$ be a weakly admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \ \textrm{ such that } \ \sqcup P_i = \cP; \   \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item The set $F(P)$ from Definition~\ref{VPropdfn} is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\item If $F$ is a cyclic flat of $M(W)$, then it is a propagator flat.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of item 1 is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. Item 2 is \cite[Lemma 3.27]{wilsonloop}. Item 3 is a direct consequence of \cite[Lemma 3.20]{wilsonloop}\sanote{ This is the right reference in both arXiv and CMP versions. Don't know what to do?} and the fact that $F(P_1)^c = V(P_1^c)$, and item 5 is given in \cite[Lemma 3.28]{wilsonloop}.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat show that certain subdiagrams of $W$ can be realized as contractions of $M(W)$, \hlfix{which we rephrase using the notation of this paper}{citation was correct, but notation was different. Karen: is this clearer?}:

\begin{lem} \label{contractsubdiaglem} \cite[Corollary 3.33]{wilsonloop} 
Let $W = (\cP, [n])$ be a Wilson loop diagram and $P \subseteq \cP$. The set $V(P)^c$ has rank $|P^c|$ (i.e. $F(P^c)$ is a propagator flat of maximal rank), if and only if the matroid defined by the subdiagram $(P, V(P))$ is equal to the contraction $M(W)/V(P)^c$.
\end{lem}
